---
title: "Analysis script to reproduce results for the 4DSW scoping review publication submitted to EEPA"
author: "Shaina Trevino"
date: "3/31/2025"
output: html_document
---

```{r renv}
# To manually load the reproducible environment
if (!require("renv")) install.packages("renv")
renv::restore()

```


```{r setup, include=FALSE}
# Install and load packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(janitor, tidyverse, rio, here, readxl, gt)

```

## Overview

This document reproduces the quantitative results reported in our publication submitted to EEPA. This file is not intended to be knitted - it is designed to be run chunk by chunk to reproduce results. The first two sections are importing and data cleaning code followed by code to reproduce our quantitative results. Code/results are presented in the order in which results are presented in the manuscript with corresponding subheadings. Data were exported on 1/29/2025 for analysis. All data files were exported directly from DistillerSR.

```{r import}
# Import study characteristics
raw_df <- import(here("data", "4dsw_study_data.xlsx")) %>% 
  janitor::clean_names() %>% 
  distinct(across(-user), .keep_all = TRUE) 

# Import eligibility data
elig_df <- read_excel(here("data", "4dsw_eligibility_data.xlsx")) %>% 
  clean_names()

# Import additional reports of studies
linked_df <- import(here("data", "4dsw_linked_references.xlsx")) %>% 
  janitor::clean_names()

# Import citations for all references
cit_df <- import(here("data", "4dsw_all_citations.xlsx")) %>% 
  janitor::clean_names()


```

```{r tidy}
# Tidy eligibility data
elig_df <- elig_df %>% 
  mutate(citation_cleaned = citation %>%
           # Remove all text between and including ##
           gsub("#[^#]*#", "", .) %>%
           # Replace multiple spaces with a single space
           gsub("\\s+", " ", .) %>%
           # Remove periods except the last one
           gsub("\\.(?![^.]*$)", "", ., perl = TRUE) %>%
           # Remove trailing spaces
           gsub("\\s+$", "", .) %>%
           # Remove instances of () and (), 
           gsub("\\(\\),\\s*", "", .) %>%
           gsub("\\(\\)", "", .) %>%
           # Replace ", ." with "."
           gsub(",\\s*\\.", ".", .)) %>% 
  #update 3/31/25 remove references added after date of data in publication
  filter(refid < 4692)

# Tidy reference data
cit_td <- cit_df %>% 
  mutate(citation_cleaned = bibliography %>%
           # Remove all text between and including ##
           gsub("#[^#]*#", "", .) %>%
           # Replace multiple spaces with a single space
           gsub("\\s+", " ", .) %>%
           # Remove periods except the last one
           gsub("\\.(?![^.]*$)", "", ., perl = TRUE) %>%
           # Remove trailing spaces
           gsub("\\s+$", "", .) %>%
           # Remove instances of () and (), 
           gsub("\\(\\),\\s*", "", .) %>%
           gsub("\\(\\)", "", .) %>%
           # Replace ", ." with "."
           gsub(",\\s*\\.", ".", .) %>% 
           # Replace " ." with "."
           gsub("\\s*\\.", ".", .) %>% 
           # Replace " ," with ","
           gsub("\\s*\\,", ",", .)) %>% 
  select(-bibliography) %>% 
  #update 3/31/25 remove references added after date of data in publication
  filter(refid < 4692)

# Transform linked references to combine multiple reports per study
linked_td <- linked_df %>% 
  select(ref_id, linked_refid) %>% 
  rowwise() %>% 
  mutate(main_refid = ref_id) %>%
  reframe(main_refid = main_refid,
          refid = c(ref_id, linked_refid)) %>%
  ungroup() %>% 
  distinct() %>% 
  left_join(cit_td) %>% 
  group_by(main_refid) %>% 
  summarize(all_reports = paste(citation_cleaned, collapse = " !!! "),
            all_refids = paste(refid, collapse = ", ")) %>% 
  ungroup() %>% 
  mutate(all_reports = str_replace_all(all_reports, "!!! ", "<br><br>")) %>% 
  rename(refid = main_refid) %>% 
  mutate(refid = as.character(refid))

# Tidy study characteristics
df_long <- raw_df %>%
  # Select relevant columns
  select(refid, starts_with("community"), starts_with("state"), starts_with("grade"),
         starts_with("school"), starts_with("effectiveness"), starts_with("equity"), 
         starts_with("evidence"), -effectiveness_approach, starts_with("stakeholder_groups_select"), starts_with("data_years_select")) %>% 
  # Pivot to long format
  pivot_longer(cols = -refid, names_to = "variable", values_to = "response") %>%
  # Extract and clean the variable categories
  mutate(
    value = sub(".*_", "", variable), # Extract the suffix of the variable name
    value = str_to_title(value), # Capitalize for consistency
    value = ifelse(value == "Unknown", "Not Reported", value), # Rename "Unknown" to "Not Reported"
    response = str_remove_all(response, "; -999|-999;"), # Clean invalid codes
    response = str_squish(response), # Remove unnecessary spaces
    # If a response exists (non-NA), use the response itself as the category
    category = case_when(
      !is.na(response) & response != "" ~ value,
      TRUE ~ NA_character_
    ),
    # Create new variable names for grouping
    new_var_name = case_when(
      str_detect(variable, "^community") ~ "community", 
      str_detect(variable, "^state") ~ "state",
      str_detect(variable, "^grade") ~ "grade_level",
      str_detect(variable, "^school") ~ "school_level",
      str_detect(variable, "^effectiveness") ~ "effectiveness",
      str_detect(variable, "^equity") ~ "equity",
      str_detect(variable, "^evidence") ~ "evidence_domain",
      str_detect(variable, "^data_years_select") ~ "data_years",
      str_detect(variable, "^stakeholder_groups_select") ~ "stakeholder_groups",
      TRUE ~ variable
    ),
    # Fix specific state naming issues
    category = case_when(
      category == "Newmexico" ~ "New Mexico",
      category == "Southdakota" ~ "South Dakota",
      category == "Northdakota" ~ "North Dakota",
      category == "Northcarolina" ~ "North Carolina",
      category == "Southcarolina" ~ "South Carolina",
      category == "Newhampshire" ~ "New Hampshire",
      TRUE ~ category
    )
  )

# Transform for calculations
td_longvar <- df_long %>%
  group_by(refid, new_var_name) %>%
  summarise(category_value = paste(category[!is.na(category)], collapse = "; "), .groups = "drop_last") %>%
  pivot_wider(names_from = new_var_name, values_from = category_value) %>% 
  ungroup() %>% 
  mutate(data_years = ifelse(data_years == "999", "Not Reported", data_years))



# List of single variables (to aggregate across id only)
single_cat_vars <- c("author_last_name", "publication_year", "publication_type", "publisher", "student_type", "student_discipline",
                     "effectiveness_approach")

# Subset of single single aggregated variables
single_df <- raw_df %>%
 select(refid, all_of(single_cat_vars))

# Transform fifth day data for table
fif_td <- raw_df %>% 
  pivot_longer(cols = fifthday_child_care:fifthday_not_reported, 
               names_to = "variable", 
               values_to = "response") %>% 
  mutate(response = str_replace_all(response, 
                                    c("Not reported" = "Not Reported"))) %>% 
  group_by(refid) %>%
  summarise(fifth_day_activities = paste(response[!is.na(response)], collapse = "; "), .groups = "drop_last") %>%
  ungroup()

# Transform additional data for table
race_td <- raw_df %>% 
  mutate(raceethnicity_checkbox_other_please_specify = str_remove_all(raceethnicity_checkbox_other_please_specify, "\\s*\\(please specify\\)")) %>% 
  pivot_longer(cols = raceethnicity_checkbox_american_indian_and_or_alaska_native:raceethnicity_checkbox_other_please_specify, 
               names_to = "variable", 
               values_to = "response") %>% 
  group_by(refid) %>%
  summarise(race_ethnicity = paste(response[!is.na(response)], collapse = "; "), .groups = "drop_last") %>%
  ungroup() %>% 
  mutate(race_ethnicity = ifelse(race_ethnicity == "", "Not Reported", race_ethnicity))

# Extract cleaned citation from elig_df
elig_subset <- elig_df %>% 
  dplyr::select(refid, citation_cleaned)

# Merge all study data together
jd <- single_df %>% 
  left_join(td_longvar) %>% 
  left_join(fif_td) %>% 
  # left_join(add_td) %>% 
  left_join(race_td) %>% 
  # left_join(ref_td) %>% 
  left_join(elig_subset) %>% 
  left_join(select(cit_df, refid, title), by = "refid")

# Correct formatting
td_study <- jd %>% 
  mutate_all(~ ifelse(is.na(.) | .x == -999 | . == "", "Not Reported", .)) %>% 
  mutate_all(~ str_remove_all(as.character(.), "-999; |; -999")) %>% 
  mutate_all(~ str_remove_all(as.character(.), "Not Reported; |; Not Reported")) %>% 
  mutate(equity = str_replace_all(equity, c("Raceethnicity" = "Race/Ethnicity", "Sexgender" = "Sex/Gender", "Ell" = "ELL", 
                                            "Ses" = "SES", "Specialeducation" = "Special Education"))) %>% 
   mutate(study_design = case_when(effectiveness_approach == "Hierarchical Linear Modeling" ~ "Between Groups - With Controls",
                                  effectiveness_approach == "Regression Adjustment" ~ "Between Groups - With Controls",
                                  effectiveness_approach == "ANOVA" ~ "Between Groups - Without Controls", 
                                  effectiveness_approach == "MANOVA" ~ "Between Groups - Without Controls", 
                                  effectiveness_approach == "Mann-Whitney U test" ~ "Between Groups - Without Controls", 
                                  effectiveness_approach == "T-Test" ~ "Between Groups - Without Controls", 
                                  effectiveness_approach == "No" ~ "Not Reported",
                                  TRUE ~ effectiveness_approach)) %>% 
  mutate(study_design = ifelse(study_design == "Not Reported", "Not Applicable", study_design),
         stakeholder_groups = ifelse(stakeholder_groups == "Not Reported", "Not Applicable", stakeholder_groups),
         outcome_domain_studied = str_replace_all(effectiveness, c("Retention" = "Staff/Teacher Retention",
                                                                   "Climate" = "School Climate",
                                                                   "Incidents" = "School Disciplinary Incidents",
                                                                   "Households" = "Household Impacts",
                                                                   "Crime" = "Criminal Activity")),
         equity_domain_studied = str_replace_all(equity, c("Age" = "Age or Grade Level",
                                                                   "Rurality" = "Community Type (Rurality)",
                                                                   "ELL" = "English Language Learner",
                                                                   "Gifted" = "Gifted Student Status",
                                                                   "Immigrant" = "Immigration Status",
                                                                   "Not Reported" = "Not Reported or Applicable")),
         study_author_year = paste(author_last_name, publication_year),
         pub_type = ifelse(publication_type == "Student", paste0(publication_type, ": ", student_type), publication_type)) 

```


## Results

### Search and Eligibility Results

This section includes code to reproduce the number of records from our search and results from eligibility screening (for PRISMA flow diagram reporting in Figure 1). The number of duplicate references removed were taken directly from DistillerSR's quarantine (from the duplicate detection feature). Bolded results are those reported in text, in addition to the PRISMA flow diagram.

Number of records identified from search & number of duplicates:

```{r num-search}
# Specify number of duplicates removed in DistillerSR
num_duplicates <- 480L

# Calculate records in full search (records screened + duplicates)
nrow(cit_td) + num_duplicates
```

Number of records screened/**Number of unique citations**:

```{r num-screen}
# Number of total records screened in DistillerSR
nrow(cit_td)
```

**Number of records excluded:** 

```{r num-drop}
# Number of records dropped at TA screening
sum(elig_df$study_screening_decision == "Drop")
```

Number of reports sought for retrieval:

```{r num-keep}
# Number of records kept at TA screening
sum(elig_df$study_screening_decision == "Keep") + 1 #Difference due to one reference that was linked to another study before TA screening was completed
```

Number of reports not retrieved:

```{r num-nopdf}
# Number of reports that we could not retrieve PDF for
sum(elig_df$pdf_retrieved == "No", na.rm = TRUE)
```

**Number of reports assessed for full-text eligibility:**

```{r num-elig-assessed}
# Number of reports that were assessed for eligibility
sum(!is.na(elig_df$study_eligibility_decision)) + 1 #Difference due to one reference that was linked to another study before TA screening was completed
```

**Number of reports excluded:**

```{r num-notelig}
# Number of reports that were not eligible during full-text assessment
sum(elig_df$study_eligibility_decision == "Not Eligible", na.rm = TRUE) + 1 #Difference due to one reference that was linked to another study before TA screening was completed
```

Exclude reasons:

```{r freq-exclude-reason}
elig_df %>% 
  filter(!is.na(study_exclude_reason)) %>% 
  count(study_exclude_reason) %>% 
  #Difference due to one reference that was linked to another study before TA screening was completed
  mutate(n = case_when(
    study_exclude_reason == "Ineligible study design - not empirical research" ~ n + 1,
    TRUE ~ n))

```

**Number of studies included in the review:**

```{r num-elig-studies}
# Number of eligible studies for data extraction
nrow(td_study)
```

**Number of total reports of included studies:**

```{r num-elig-reports}
# Number of reports that were eligible during full-text assessment
sum(elig_df$study_eligibility_decision == "Eligible", na.rm = TRUE)
```

### Characteristics of Sources of Evidence

```{r tbl-results}
# Create functions to calculate numbers reported in Table 1 and 2
# Transform data for table
data <- raw_df %>% 
  mutate(across(where(is.character), ~ na_if(., "")))

# Define function to summarize checkbox variables
summarize_checkboxes <- function(data, prefix, filter_condition = NULL) {
  filtered_data <- if (!is.null(filter_condition)) {
    data %>% filter(!!rlang::parse_expr(filter_condition))
  } else {
    data
  }
  
  filtered_data %>%
    select(starts_with(prefix)) %>%
    summarise(across(everything(), ~ sum(!is.na(.)), .names = "N_{col}")) %>%
    pivot_longer(cols = everything(), names_to = "Category", values_to = "N") %>%
    mutate(
      Percentage = round((N / nrow(filtered_data)) * 100, 1),
      Denominator = nrow(filtered_data)  # Add the denominator
    )
}
  
# Apply function to variable groups
community_summary <- summarize_checkboxes(data, "community_")
state_summary <- summarize_checkboxes(data, "state_")
grade_summary <- summarize_checkboxes(data, "grade_")
school_summary <- summarize_checkboxes(data, "school_")
fifthday_summary <- summarize_checkboxes(data, "fifthday_")
evidence_summary <- summarize_checkboxes(data, "evidence_")
outcome_summary <- summarize_checkboxes(data, "effectiveness_", filter_condition = "!is.na(evidence_effectiveness)") #demon = studies reporting effectiveness
equity_summary <- summarize_checkboxes(data, "equity_", filter_condition = "!is.na(evidence_equity)") #demon = studies reporting equity
stakeholder_summary <- summarize_checkboxes(data, "stakeholder_groups_select_", filter_condition = "!is.na(evidence_acceptability)") #demon = studies reporting acceptability



# Create function to summarize single-response categorical variables
categorical_summary <- function(data, var, filter_condition = NULL) {
  filtered_data <- if (!is.null(filter_condition)) {
    data %>% filter(!!rlang::parse_expr(filter_condition))
  } else {
    data
  }
  
  filtered_data %>%
    count(!!sym(var)) %>%
    mutate(
      Percentage = round((n / nrow(filtered_data)) * 100, 1),
      Denominator = nrow(filtered_data)  
    ) %>%
    rename(Category = !!sym(var), N = n)
}
 
# Apply function
publication_summary <- categorical_summary(data, "publication_type")
student_type_summary <- categorical_summary(data, "student_type", filter_condition = "publication_type == 'Student'")
student_disc_summary <- categorical_summary(data, "student_discipline", filter_condition = "publication_type == 'Student'")
race_summary <- categorical_summary(data, "raceethnicity_majority")
approach_summary <- categorical_summary(data, "effectiveness_approach", filter_condition = "!is.na(evidence_effectiveness)")


# Create function to summarize numerical variables (median and range)
numerical_summary <- function(data, var) {
  data %>%
    summarise(
      Median = median(!!sym(var), na.rm = TRUE),
      Range = paste0(min(!!sym(var), na.rm = TRUE), "-", max(!!sym(var), na.rm = TRUE))
    )
}

# Apply function
publication_year_summary <- numerical_summary(data, "publication_year")

# For data_years, as checkbox variable
# Combine all year columns into a single column
years_summary <- data %>%
  select(starts_with("data_years_select_")) %>%  # Select all year-related columns
  pivot_longer(cols = everything(), names_to = "year_variable", values_to = "year") %>%  # Reshape to long format
  mutate(year = as.numeric(year)) %>%  # Ensure the years are numeric
  filter(!is.na(year) & year != -999)  # Remove NA and -999 values

# Calculate median and range
summary_data_years <- years_summary %>%
  summarise(
    Median = median(year, na.rm = TRUE),
    Range = paste0(min(year, na.rm = TRUE), "-", max(year, na.rm = TRUE))
  )


```

Dates included studies were published:

```{r sum-year}
# Median and Range for publication year of included studies
print(publication_year_summary)
```

Publication types of included studies:

```{r sum-pubtype}
# Frequency table for type of publication
print(publication_summary)
```

Student publication types: 

```{r sum-studtype}
# Frequency table for student publication types
print(student_type_summary)
```

Student disciplines for student publications:

```{r sum-studdis}
# Frequency table for student publication disciplines
print(student_disc_summary)
```

Dates studies collected and analyzed data:

```{r sum-datayrs}
# Median and Range for data years
print(summary_data_years)
```

Community types among included studies:

```{r sum-rurality}
# Frequency table for community type
print(community_summary)
```

States included:

```{r sum-states}
# Frequency table for all states
print(state_summary, n = 50) %>% arrange(desc(N))
```

School grade level included:

```{r sum-schl}
# Frequency table for school grade level
print(school_summary)
```

Student grades included:

```{r sum-grade}
# Frequency table for student grade level
print(grade_summary)
```

Student race and ethnicity:

```{r sum-race}
# Frequency table for student race/ethnicity
print(race_summary)
```

Fifth day activity offered by schools:

```{r sum-fd}
# Frequency table for fifth day activity
print(fifthday_summary)
```

### Types of Empirical Research Evidence

Number of studies reporting information on effectiveness: 

```{r num-effectiveness}
# Number of studies reporting effectiveness
print(transform(outcome_summary[1,], Percentage = round(N / nrow(data) * 100, 0), Denominator = nrow(data))) 
```

Among effectiveness studies, methodological approaches used:

```{r sum-design}
# Frequency table of study design
print(approach_summary) %>% arrange(desc(N))
```

Among effectiveness studies, outcome domains:

```{r sum-outcome}
# Frequency table of outcome domains
print(outcome_summary) %>% filter(Category != "N_effectiveness_approach") %>% arrange(desc(N))
```

Among effectiveness studies, number reporting equity:

```{r num-equity}
# Number of effectiveness studies reporting equity domain
print(transform(evidence_summary[3,], Percentage = round(N / nrow(data %>% filter(!is.na(evidence_effectiveness))) * 100, 0))) 
```

Among equity studies, equity domains:

```{r sum-equity}
# Frequency table for equity domains
print(equity_summary)
```

Number of studies reporting info on implementation: 

```{r num-implementation}
# Number of studies reporting any info on acceptability, feasibility, or resource use
n_implementability <- data %>%
  filter(
    !is.na(evidence_acceptability) |
    !is.na(evidence_feasibility) |
    !is.na(evidence_resource)
  ) %>%
  distinct(refid) %>%
  nrow()

paste0(n_implementability, " (", round((n_implementability / nrow(data)) * 100, 1), "%) ", "studies reported implementability")

```

Among studies reporting implementation, number of studies reporting evidence on acceptability: 

```{r num-acceptability}
# Number of studies reporting acceptability (from studies reporting implementation)
print(transform(evidence_summary[1,], Percentage = round(N / n_implementability * 100, 0), Denominator = n_implementability)) 
```

Among acceptability studies, target of acceptability:

```{r sum-acceptability}
# Frequency of stakeholder groups targeted for acceptability domain
print(arrange(stakeholder_summary, desc(N)), n = 30) 
```

Among studies reporting implementation, number of studies reporting evidence on feasibility: 

```{r num-feasibility}
# Number of studies reporting feasibility (from studies reporting implementation)
print(transform(evidence_summary[4,], Percentage = round(N / n_implementability * 100, 0), Denominator = n_implementability)) 
```

Among studies reporting implementation, number of studies reporting evidence on resource use: 

```{r num-resource}
# Number of studies reporting resource use (from studies reporting implementation)
print(transform(evidence_summary[6,], Percentage = round(N / n_implementability * 100, 0), Denominator = n_implementability)) 
```

Number of studies reporting a policy priority: 

```{r num-priority}
# Number of studies reporting policy priority (out of all studies)
print(evidence_summary[5,])
```

## Figures

### Figure 1. Flow Diagram for Selection of Sources of Evidence

Figure 1 was created in word to follow the PRISMA 2020 reporting template. Code to reproduce the numbers reported in Figure 1 can be found in the "Search and Eligibility Results" section above.  

## Tables

The code below will reproduce the values we reported within tables in the manuscript. These results will be generated in the console - no tables will be generated or saved as outputs. 

### Table 1. Characteristics of Included Sources of Evidence

Results for Table 1

```{r print-tbl1}
# Print Table 1 info
print(publication_summary)
print(community_summary)
print(school_summary)
print(race_summary)
print(fifthday_summary)

# Table note
print(publication_year_summary)
print(summary_data_years)

```


### Table 2. Available Empirical Research Evidence on Four-Day School Weeks

Results for Table 2

```{r print-tbl2}
# Print Table 2 info
evidence_summary %>%
  mutate(Denominator = if_else(Category == "N_evidence_equity",
                               N[Category == "N_evidence_effectiveness"],
                               Denominator),
         Percentage = if_else(Category == "N_evidence_equity",
                              round(N / Denominator * 100, 1),
                              Percentage))

```


## Supplements

The code below will reproduce the values we reported in supplemental tables. These results will be generated in the console - no tables will be generated or saved as outputs. 

### Online Supplement 1. Search Strategy

There is no code to produce S1 since it was created in Word to list our search strategy.

### Online Supplement 2. Characteristics of Included Sources of Evidence 

Results for S2

```{r print-supp1}
# Print S2 info
print(publication_summary)
print(student_type_summary)
print(student_disc_summary) 
print(community_summary)
print(state_summary, n = 50)
print(grade_summary)
print(school_summary)
print(race_summary)
print(fifthday_summary)

# Table note
print(publication_year_summary)
print(summary_data_years)

```

### Online Supplement 3. Available Empirical Research Evidence on Four-Day School Weeks 

Results for S3 

```{r print-supp2}
# Print Table 2 info
print(evidence_summary[1,])
print(stakeholder_summary, n = 30) 
print(evidence_summary[2,])
print(approach_summary) 
print(outcome_summary) 
print(transform(evidence_summary[3,], Percentage = round(N / nrow(data %>% filter(!is.na(evidence_effectiveness))) * 100, 0), Denominator = nrow(data %>% filter(!is.na(evidence_effectiveness))))) 
print(equity_summary)
print(slice(evidence_summary, -1:-3))

```

